---
title: "uas visualisation"
author: "Mukul Chaware"
date: "5 April 2016"
output: html_document
---
```{r,eval=T,warings=F,message=F}
##loading libraries
library(ggplot2)
library(ggthemes)
library(readxl)
library(dplyr)
library(grid)
library(ggmap)
library(gganimate)
library(animation)
```
```{r,eval=T,warnings=F,message=F}
xl1 <- read_excel("UAS_Sightings_report_21Aug-31Jan.xlsx")
xl2 <- read_excel("UASEventsNov2014-Aug2015.xls")
drones=read.csv("drones.csv")
drones$month = factor(drones$month,levels(drones$month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])

drones2=read.csv("drones2.csv")
drones2$month = factor(drones2$month,levels(drones$month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])

by_state <- drones2[drones2$class=="NOTIFIED",] %>%
  group_by(state) %>%
  summarize(count = n()) %>%  mutate(region = tolower(state))

missing_states <- data_frame(
  region = c("wyoming", "vermont", "nebraska", "iowa","idaho","north dakota","south dakota"),
  count=c(0,0,0,0,0,0,0)
)
by_state <- bind_rows(by_state, missing_states)

us <- map_data("state")

mid <- mean(by_state$count, na.rm = TRUE) 
```

```{r,eval=T,warning=F,message=F}
p=ggplot(data=drones) + 
 geom_bar(mapping=aes(x=hour,fill="red"))+ggtitle("Hourly classification of UAS")+ylab("# UAS sightings")+ scale_fill_discrete("",labels="",breaks="")
plot(p)
```

Surprsingly count is very high at midnight (near 12 AM) and very low in early morining.

```{r,eval=T,warning=F,message=F}
p=ggplot(data = drones[!is.na(drones$month),]) + 
  geom_bar(mapping = aes(x = month,fill=hour_class),position="fill")+ggtitle("Classify Monthly UAS sightings on the basis of day and night time")+xlab("# UAS sightings")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot(p)
```

I have considered 7 AM to 6PM as Day time and 7 Pm to 6 AM as Night time. The reason sightings in august and September is very high in night time due to spring (that's my guess). 

```{r,eval=T,warning=F,message=F,fig.show='animate'}
p=ggplot() + 
  geom_map(aes(x = long, y = lat, map_id = region), data = us,
           map = us,fill="gray",color="white") + 
  geom_point(mapping = aes(x=lon,y=lat,color=hour_class,frame=month),data=subset(drones2,lon>= min(us$long) & lon<=max(us$long) &   lat>= min(us$lat) & lat<=max(us$lat)  ),size=5,alpha=0.7)+
  ggtitle("UAS occurences month : ")+theme_light()

ani.options(interval = 1, ani.width = 800, ani.height = 800)
gg_animate(p,"monthly UAS occurences.gif")
```
```{r,eval=T,warning=F,message=F}
p=ggplot(data = drones2) + 
  geom_bar(mapping = aes(x = class,fill=class))+ggtitle("Whether UAS was notified to LEO or not")
plot(p)
```

Large number of cases are notified to nearby LEOs. Very few cases are not notified. 

```{r,eval=T,warning=F,message=F}
####UASEventsNov2014-Aug2015
p=ggplot(data = drones2[1:nrow(xl2),]) + 
  geom_bar(mapping = aes(x = class,fill=class)) +ggtitle("UASEventsNov2014-Aug2015")
plot(p)
```
```{r,eval=T,warning=F,message=F}
p=ggplot(data = drones2[(nrow(xl2)+1):nrow(drones2),]) + 
  geom_bar(mapping = aes(x = class,fill=class)) +ggtitle("UAS_Sightings_report_21Aug-31Jan")
plot(p)
```
```{r,eval=T,warning=F,message=F}
p=ggplot(data=drones2) +
  geom_bar(mapping = aes(x=state,fill=class))+ ggtitle("Stateweise classification")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot(p)
```

California has very high number of drone sightings.Unknown cases (don't know whther they are notified or not) and not notified cases are also high in california. Other than California, Florida and New York has high drone sightings. 

```{r,eval=T,warning=F,message=F}

p=ggplot(data=drones2[drones2$yw>=201500 & drones2$yw<=201570,]) + 
  geom_bar(mapping=aes(x=yw,fill=class))+ggtitle("Weekly classification in year 2015")
plot(p)
```

```{r,eval=T,warning=F,message=F}
p=ggplot() +
  geom_map(aes(x = long, y = lat, map_id = region), data = us,
           map = us, fill = "#ffffff", color = "#ffffff", size = 0.15) +
  geom_map(aes(fill = count, map_id = region),
           map = us, data = by_state,
           color = "#ffffff", size = 0.15)+ggtitle("UAS map")
plot(p)
```
```{r,eval=T,warning=F,message=F,fig.show='animate'}
p=ggplot() + 
geom_map(aes(x = long, y = lat, map_id = region), data = us,
         map = us,fill="gray",color="white") + 
    geom_point(mapping = aes(x=lon,y=lat,color=class,frame=yw),data=subset(drones2,lon>= min(us$long) & lon<=max(us$long) &   lat>= min(us$lat) & lat<=max(us$lat)  ),size=7)+theme_light()+ggtitle("Weekly UAS occurences")

ani.options(interval = 0.5, ani.width = 800, ani.height = 800)
gg_animate(p,"weekly UAS occurences.gif")
```

